\documentclass[../Main.tex]{subfiles}

\begin{document}

Certificate of Need (CON) laws impose bureaucratic costs by requiring healthcare providers who plan on opening a new healthcare facility, expanding a current facility, or purchasing medical equipment to first show to a regulatory body that their region needs the service. Nursing home CON regulations have historically been the most prevalent CON regulation. Figure \ref{fig:nh_histogram} shows that 36 American states regulate nursing homes via CON laws, which is more than any other health service. Often, CON regulations on nursing homes are also more restrictive than other health services. While all CON regulations impose a bureaucratic cost of applying for a license, it is still possible to get the license and offer the new service for most categories of CON regulations. Nursing home CON, however, has been applied in many states as a moratorium. Even states that do not have a moratorium still require nursing homes to go through a long and costly review process to get a permit for any additional beds \citep{american2020american}. Nursing home CON laws have also historically been the first CON regulation to be adopted and the last to be repealed. Ohio, for example, dropped all its CON regulations in 1997 except for its CON regulation on nursing homes. Indiana revoked all its CON regulations in 1999 but reinstated only nursing home CON in 2019. Thus, nursing homes are heavily regulated by CON laws to an extent that no other health service is. 
 
  %\footnote{In this paper, the term nursing home (often called skilled nursing homes or long-term care facilities) refers only to in-patient rehabilitation and medical treatment centers staffed with trained medical professionals. Restriction levels in this paper are defined as the additional cost that firms have to incur to provide a nursing home service. These are often associated with high fees, ``red-tape,'' and a low probability of receiving authorization to offer the service. Thus, an increase in restrictions implies that the cost of applying for a CON went up or that the probability of receiving the CON went down.} 

Nursing home services have historically represented over a third of Medicaid expenditures \citep{wiener1999controlling}, with Medicaid being the primary payer for nursing home services, covering about two thirds of all nursing home residents \defcitealias{ahca2020pressrelease}{AHCA/NCAL,~2020}\citepalias{ahca2020pressrelease}. The connection between nursing homes and Medicaid is often cited by politicians and healthcare executives as the primary reason to regulate nursing homes via CON. Emmett Reed, Florida Health Care Association’s executive director, argued for example, that the repeal of nursing home CON in Indiana led to overproduction in the development of new nursing facilities which put a strain on the state’s Medicaid budget, ultimately leading lawmakers to backtrack and reinstate its CON law in 2019. He further concluded that nursing home CON must be maintained because “the burden [of repealing CON] ultimately will fall to the taxpayers” \citep{sexton2019conrepeal}. With its presumed effect on Medicaid expenditures, many regulators may view nursing home CON as one of, if not the most, necessary CON restriction.

This paper explores why nursing home CON regulations are so prevalent, and the effect that such regulations have on the quantity and quality of, and expenditures on, nursing home services. We develop a dynamic model in which state politicians choose their level of nursing home CON regulation; nursing homes choose their level of quality, after observing the regulators of all states; and consumers choose their nursing homes, after observing the quality of all nursing homes. The model predicts that implementing (repealing) this regulation leads to a decrease (an increase) in nursing home quality, quantity, and Medicaid expenditures on nursing home services. The model also shows that interstate competition to avoid nursing home Medicaid costs creates incentives for state politicians to impose nursing home CON restrictions higher than they would impose in the absence of such competition. 

Our model predictions and theoretical framework suggest that the decision of state politicians to regulate nursing homes through CON are analogous to decisions over taxing capital investments and offering welfare benefits. If a state taxes capital investments at lower rates than its neighbors, more capital investments will be made in this state, all else equal \citep{zodrow1986pigou,bucovetsky1991asymmetric, kanbur1993jeux, basinger2004remodeling, plumper2009there}. If a state offers more welfare transfers than other states, more people who would benefit from such transfers will move to that state, all else equal \citep{gramlich1984migration, peterson1989american, saavedra2000model}. Likewise, we find that if a state imposes lower nursing home CON restrictions than other states, more nursing home consumers will move to that state, all else equal. Therefore, repealing or reducing nursing home CON restrictions is expected to cause an increase in Medicaid expenditures.

All predictions from our model are specific to the nursing home market. Nursing homes are unique among health services because individuals in need of nursing home services often move across states to get these services. This is especially true for consumers who live near state borders or who have family members who live in other states. Consumers of other health services may travel across states to get these services but rarely move because of it. This distinction is essential because Medicaid expenditures are tied to one's residency. When a nursing home consumer who is eligible for Medicaid moves to a new state, they have to update their residency and the new state pays for their Medicaid expenses.  

We assess the predictions of our model empirically by analyzing the repeal of nursing home CON regulations in Pennsylvania (PA), Indiana (IN), and North Dakota (ND). More specifically, we use the synthetic difference-in-differences method \citep{arkhangelsky2021synthetic} to estimate the effect in each of these states of removing nursing home CON regulations on the quantity and quality of nursing home services (as measured by the quantity of nursing home beds, nursing homes, and specialized care beds per 100,000), as well as on Medicaid nursing home expenditure per capita. The synthetic difference-in-differences estimator builds on the insights behind the widely used difference-in-differences and synthetic control methods \citep{abadie2003economic,abadie2010synthetic,currie2020technology}, and is argued by \citet{arkhangelsky2021synthetic} to perform well in settings such as ours, where these conventional estimators are commonly used. We use the synthetic difference-in-differences method because \citet{arkhangelsky2021synthetic} show that this estimator has desirable robustness properties relative to synthetic control and difference-in-difference methods. Like synthetic control, the synthetic difference-in-differences method reweights and matches pre-exposure trends in order to weaken the reliance on parallel trend type assumptions. Synthetic difference-in-differences also includes time weights that lessen the role of time periods that are very different from the post-treatment periods. Like differences-in-differences, the synthetic difference-in-differences method is invariant to additive unit-level shifts, and allows for valid large-panel inference.

We find that the repeal of nursing home CON regulations in PA and IN caused nursing home Medicaid expenditure per capita to increase by about 70\% and 38\%, respectively, relative to pre-repeal levels of expenditure. These represent increases in total annual nursing home Medicaid expenditure of roughly \$1.32 billion in PA, and \$334 million in IN. Unfortunately, the estimated effects of the repeal on Medicaid expenditure in ND are unreliable due to a lack of pre-treatment parallel trends. Consistent with the predictions of our model, we find that the increases in Medicaid expenditure in PA and IN are being driven at least in part by increases in the quantity of nursing home services. According to our estimates, repealing nursing home CON caused an increase in both the quantity of nursing home beds and nursing home facilities per 100,000 in PA and IN, with the effects being relatively larger in PA. These findings imply that facilities in treated states soon after the repeal were much newer than in their counterfactual, which can factor into consumer's perceptions of nursing home quality. We find that repealing nursing home CON also caused an increase in specialized care beds per 100,000, again most notably in PA, which \citet{grabowski2010quality} show is correlated with higher quality of service. 

We also analyze the effect of repealing nursing home CON on total nursing home expenditure per capita, and find that it is very similar to the estimated effect on nursing home Medicaid expenditure per capita. This suggests that most of the effect of repealing CON on total expenditure is likely coming from its effect on Medicaid expenditure. We further analyze this hypothesis by testing if the repeal had any effect on Medicare expenditure and find no strong evidence that it did. If the estimated increase in Medicaid nursing home expenditure is in fact being caused by the repeal of nursing home CON, as opposed to some other factor, then we would not expect to see a corresponding increase in less related types of Medicaid expenditure, such as hospital and home care Medicaid expenditure. We find no significant effects of repealing nursing home CON on Medicaid hospital and home care expenditure per capita in PA, IN, or ND. 

All results where we find evidence for parallel trends are consistent with the predictions of our model. In particular, we find that repealing nursing home CON causes an increase in the quantity and quality of nursing home services, as well as an increase in Medicaid expenditure on those services. Moreover, our empirical results show that the effects of repealing nursing home CON can be substantial and economically meaningful. Our findings are also consistent with the proposition that inter-state competition within the current Medicaid system incentivizes state politicians to impose nursing home CON regulations that are more restrictive than they would be in the absence of such inter-state competition.

We contribute to the literature that explores how different policies affect Medicaid expenditure, and in particular Medicaid expenditure on nursing homes. \citet{grabowski2004recent} show how the repeal of the Boren amendment, which gave states greater freedom to set Medicaid nursing home policies, affected Medicaid expenditures. \citet{goda2011impact} analyzes the effects of subsidies to private long-term care insurance on Medicaid expenditures. \citet{grabowski2003effects} use a two-way fixed effects analysis and find that nursing home CON did not have a significant effect on Medicaid expenditures. \citet{rahman2016impact} exclude states that changed their CON regulations in their period of study and find that states with nursing home CON laws had faster Medicaid and Medicare expenditure growth than states without CON laws. \citet{bailey2019can} finds that the presence of any CON regulation (not necessarily nursing home CON) in a state is positively associated with total health expenditures and total nursing home expenditure, which is opposite to the effects predicted by our model and supported by our empirical analysis. However, Bailey’s specification analyzes the impact of the presence of any CON regulation as opposed to nursing home CON in particular. Many have also analyzed the correlation between nursing home CON and the quantity of nursing home services. \citet{harrington1997effect}, \citet{swan1991certificate}, and \citet{zinn1994market} for example, find that the presence of nursing home CON laws are associated with reduced growth in the number of nursing home beds. We contribute to this literature mainly by constructing valid causal estimates of the (potentially heterogeneous) effects of a state repealing nursing home CON on the quantity and quality of nursing home services, as well as on nursing home Medicaid expenditure.   

While prior research has explored questions of competitive federalism, we offer the first political economy analysis of nursing home CON laws and the first inter-jurisdictional competition model applied to the context of healthcare policy. \citet{tiebout1956pure} was the first to show that competition across jurisdictions places competitive pressures on local governments that lead to predictable policy choices. He shows that, under certain conditions, competitive pressures lead governments to provide the optimal level of public goods. \citet{zodrow1986pigou} formalize the mechanisms of what became known as tax competition models. They model two countries competing for a perfectly mobile capital stock and show that, in equilibrium, tax rates are lower in both countries than they would be otherwise. \citet{bucovetsky1991asymmetric} and \citet{kanbur1993jeux} show that differences in country size can substantially change the predictions of the model such that smaller countries face stronger incentives to reduce taxes. \citet{genschel2002globalization} explore how institutional constraints affect policy adjustment to competitive pressures. \citet{basinger2004remodeling} and \citet{plumper2009there} show that political costs mitigate the competitive downward pressure on tax rates. 

A large literature similarly explores the effects of inter-state competition on political decisions over welfare policies. \citet{gramlich1984migration} model the effect of migration on state welfare policy and find that competition across jurisdictions incentivizes political representatives to provide lower welfare benefits than they otherwise would. \citet{peterson1989american} explores the same question and develops the magnet hypothesis which proposes that an increase in state welfare benefits may lead to an increase in poverty rates because the state becomes a magnet that attracts some low-income citizens from other states. Our model predicts that the repeal of nursing home CON will lead to an increase in the number of nursing homes services and Medicaid expenditures because states will become a ``magnet'' for nursing home consumers. This proposition is analogous to the welfare magnet hypothesis proposed by \citet{peterson1989american}. \citet{saavedra2000model} develops a model analogous to the ones from the tax competition literature and shows that state choices of welfare benefits face a downward pressure from interstate competition. We contribute to the literature on competitive federalism by modeling how migration associated with the consumption of nursing home services leads to inter-state competition over nursing home CON restrictions. 

%We adapt the structure of the tax competition model from \citet{basinger2004remodeling} to the context of nursing home CON regulation. Our model predicts that the repeal of nursing home CON will lead to an increase in the number of nursing homes services and Medicaid expenditures because states will become a ``magnet" for nursing home consumers. This proposition is analogous to the welfare magnet hypothesis proposed by \citet{peterson1989american}. In the absence of political marginal costs, we show that equilibrium nursing home CON regulations spiral upward leading to a race to the top in health regulation. Once we consider political constraints, we find that political costs can mitigate this competitive upward pressure. Most importantly, the model shows that inter-state competition can incentivize state politicians to impose nursing home CON restrictions that are higher than they would be in the absence of such competition.

%The paper is organized as follows: Section \ref{policy_background} overviews the policy background of CON laws and its literature with a focus on nursing homes; Section \ref{model} describes the model used to understand the impact of nursing home CON on access to nursing home services and nursing home expenditures; Section \ref{data} describes the data sources used; Section \ref{empirical_strategy} summarizes the synthetic difference in differences method and contrasts this empirical strategy with the traditional synthetic control and difference in differences methods; Section \ref{results} highlights the results of the synthetic difference-in-difference analysis and also contrasts it with the results from the synthetic control and difference in differences analyses; Section \ref{conclusion} offers a conclusion and the policy implications of this research.


\end{document}