\documentclass[../Main.tex]{subfiles}

\begin{document}

NH-CON laws impose bureaucratic costs to nursing home providers by requiring those planning on opening a new nursing home or expanding a current facility to first show to a regulatory body that their region needs the service. In the context of a simple supply and demand model, these regulations would act to decrease the supply of nursing home services. Relative to an unregulated industry, therefore, we would expect to see a higher price, a lower quantity, and excess demand involving individuals willing to pay the unregulated price, but not the regulated price.\footnote{An alternative way to think about NH-CON laws, especially the most restrictive regulations involving a moratorium, is that they impose a restriction on the quantity of nursing home services/beds. This restriction would result in a higher price, a lower quantity, and excess demand, the same outcome as modelling the regulations as a government-imposed decrease in supply.} It follows that when PA, IN, and ND dropped their NH-CON regulations, a resultant decrease in price and increase in quantity would be expected.\\ 
\indent When PA, IN, and ND dropped their NH-CON laws, this likely not only affected the incentives of producers and consumers in Pennsylvania, but also the incentives of individuals in surrounding states. The reason is that many states surrounding PA, IN, and ND (and in the case of PA, all surrounding states) maintained their NH-CON regulations when PA, IN, and ND dropped theirs, and were therefore in a similar position with excess demand at the prices and quantities resulting from the regulation. As prices dropped and quantities of nursing home services/beds increased as a result of dropping NH-CON in PA, IN, and ND, it is likely that at least some elderly individuals in the surrounding states would migrate to the new nursing home facilities. This increase in demand from surrounding states would act to increase prices and further increase quantities.\\ 
\indent Taking the direct effects in PA, IN, and ND of dropping their CON regulations, together with the indirect effects from surrounding states, we would expect to see an unambiguous increase in the quantity of nursing home services in PA, IN, and ND. The effect on government expenditure, however, which is a function of both price and quantity, is theoretically ambiguous. That notwithstanding, if there is a significant increase in demand from surrounding states when NH-CON laws are dropped, the resulting combination of higher prices and quantities would likely result in an overall increase in expenditure, particularly Medicaid expenditure since Medicaid, not Medicare, covers long-term care. Furthermore, since Medicaid is a joint federal and state program, individuals moving from surrounding states wishing to utilize Medicaid to finance their nursing home services in PA, IN, or ND would need to change their residency status, and would therefore start receiving Medicaid benefits from PA, IN, or ND. Thus, by dropping their CON laws, PA, IN, and ND could end up paying part of the bills for individuals from surrounding states.\\
\indent The in-migration from surrounding NH-CON states depends in large part on how good of a substitute the nursing home services are in the states that dropped NH-CON regulations. We argue that of the three states that dropped NH-CON regulations, PA provides the best substitute for nursing home services since it is near large urban centers, has similar/warmer weather to its neighbors, and has major cities near large metropolises. Therefore, our hypothesis is that the quantity of nursing home services would increase in all three states, but total government and Medicaid nursing home expenditure would only increase, or at least would increase the most, in PA compared to in IN and ND.\\
\indent Our findings are consistent with this hypothesis. Figures \ref{fig:q_nursing_homes_trends} and \ref{fig:q_nursing_home_beds_trends} show trends in the quantity of nursing homes and nursing home beds per 100,000, respectively, in PA, IN, and ND and their corresponding synthetic controls. The quantity of nursing homes and beds per 100,000 in PA, IN, and ND and their corresponding synthetic controls are virtually identical up until they dropped NH-CON regulations, at which point there is a clear divergence with the quantities in the states that dropped NH-CON increasing relative to their synthetic controls. Recall from Section \ref{empirical_strategy} that we only used the first half of the pre-intervention years to form the synthetic matches using our outcome variable, and reserved the second half for out-of-sample validation. Despite only matching on the first half of the pre-intervention period, the outcomes in each state match their synthetic control extremely well in the second half of the pre-intervention period as well, instilling confidence that the synthetic controls replicate well the respective counterfactual trajectories of PA, IN, and ND in the post-intervention period.\\
\indent The year-specific effects of dropping NH-CON regulations on the quantity of nursing homes and beds, $\hat{\alpha}_{it}$ from equation (\ref{eq:year_spec_effect}), are shown graphically in Figures \ref{fig:q_nursing_homes_gaps} and \ref{fig:q_nursing_home_beds_gaps}, respectively. Also shown in these figures are the year-specific placebo effects for the states in the donor pool of potential control states, allowing for a visual inspection of how large the year-specific effects are in the treated states in comparison to the distribution of placebo effects. Our findings indicate that dropping NH-CON caused an immediate increase in the quantity of nursing homes per 100,000 in all three states with a subsequent gradual increase in PA and IN, and levelling off in ND. While there is less of an initial immediate increase, especially in PA and IN, a similar pattern emerges for the quantity of nursing home beds as shown in Figure \ref{fig:q_nursing_home_beds_gaps}. The year-specific effects in the treated states, especially PA and IN, appear to be relatively large when compared to the the distribution of year-specific placebo effects.\\
\indent Columns (3) and (4) of Table \ref{tab:ave_results} report the average annual effects, $\bar{\alpha}_1$ from equation (\ref{eq:ave_effect}), of dropping NH-CON regulations on the quantity of nursing homes and beds per 100,000, respectively. Dropping NH-CON caused an increase of 0.37, 0.28, and 0.21 nursing homes per 100,000 in PA, IN, and ND, respectively. The exact p-values corresponding to these estimates are 0.083 for PA, 0.028 for IN, and 0.472 for ND, suggesting that only the estimates in PA and IN are statistically significant at any reasonable significance level. These estimated effects are quite large when compared to the average quantity of nursing homes per 100,000 over the pre-intervention period. For example, an increase of 0.37 and 0.28 nursing homes in PA and IN is an increase of about 58.7 and 44.4 percent relative to their pre-intervention averages of 0.63. We find a similar positive effect on the quantity of nursing home beds per 100,000 in all three states, with the effect being largest and statistically significant in IN.\\
\indent Figures \ref{fig:tot_exp_trends} and \ref{fig:medicaid_exp_trends} show trends in total government nursing home and Medicaid expenditure per capita, respectively. Similar to the quantity trends, we find that our synthetic controls do an extremely good job of matching their corresponding treated states (with the exception of ND in Panel (C) of Figure \ref{fig:medicaid_exp_trends}), despite only matching in the first half of the pre-intervention period. There's a clear divergence in total and Medicaid expenditure per capita after NH-CON is dropped only in PA, with total expenditure growing much faster in PA than in its synthetic control, and Medicaid expenditure increasing dramatically, especially initially, while Medicaid expenditure would have remained relatively stagnant had it not dropped NH-CON. Figures \ref{fig:tot_exp_gaps} and \ref{fig:medicaid_exp_gaps}, which show the year-specific effects in PA, IN, and ND and the placebo effects of dropping NH-CON on total and Medicaid nursing home expenditure, tell  the same story, and suggest that total and Medicaid expenditure increased dramatically only in PA relative to the distribution of placebo effects.\\
\indent Columns (1) and (2) of Table \ref{tab:ave_results} report the average annual effects of dropping NH-CON on total and Medicaid nursing home expenditure per capita. Interestingly, the estimates of the effect on Medicaid expenditure are negative in both IN and ND, and also on total expenditure in IN. The exact p-values corresponding to each of these estimates, however, are relatively large suggesting no statistically significant effect of dropping NH-CON on total or Medicaid expenditure in IN or ND. In contrast, we find large and statistically significant effects of dropping NH-CON on total and Medicaid expenditure per capita in PA. We estimate that dropping NH-CON caused total annual expenditure to increase by \$123.57 per capita, an increase of about 35.3 percent relative to the average pre-intervention total expenditure of \$361 per capita. Most of this increase in total nursing home expenditure was in the form of increased nursing home Medicaid expenditure. We estimate that dropping NH-CON caused nursing home Medicaid expenditure to increase by \$104.17 per capita, an increase of about 71.6 percent relative to the average pre-intervention nursing home Medicaid expenditure of \$145.56 per capita. These results are consistent with a relatively large increase in demand coming from states surrounding PA for it's good substitute nursing home services, suggesting that by dropping its NH-CON regulations, PA ended up paying a considerable amount in Medicaid expenditure for individuals who would have otherwise been residents, and therefore Medicaid recipients, of surrounding states.

\end{document}